News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-09-04 rupnik2015.tex
related work

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CL-LSI\shortcite{cl_lsi}\cite{multilingualBook}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CL-LSI. Another closely related method is Explicit Semantic Analysis (ESA) \cite{ESA}, which uses Wikipedia (as do we in the current work) to compare documents. The authors of ESA compare it to CL-LSI and find that CL-LSI can outperform ESA in an information retrieval, but is costlier to optimize over a large corpus (ESA requires no training). We find that the scalability argument does not apply in our case: based on advances in numerical linear algebra we can solve large CL-LSI problems (millions of documents as oposed to 10,000 document limit reported in \cite{ESA}).

@@

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$. 
When building cross-lingual models, the IDF scores were computed with respect to the Wikipedia corpus. In the other part of our system, we computed TFIDF vectors on streams of news documents in multiple languages. There the IDF scores for each language changed dynamically - for each new document we computed the IDF of all news articles within a 10 day window.

>>>>

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CL-LSI\shortcite{cl_lsi}\cite{multilingualBook}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CL-LSI. Another closely related method is Explicit Semantic Analysis (ESA) \cite{ESA}, which uses Wikipedia (as do we in the current work) to compare documents. The authors of ESA compare it to CL-LSI and find that CL-LSI can outperform ESA in an information retrieval, but is costlier to optimize over a large corpus (ESA requires no training). We find that the scalability argument does not apply in our case: based on advances in numerical linear algebra we can solve large CL-LSI problems (millions of documents as opposed to 10,000 document limit reported in \cite{ESA}).

@@

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$. 
When building cross-lingual models, the IDF scores were computed with respect to the Wikipedia corpus. In the other part of our system, we computed TFIDF vectors on streams of news articles in multiple languages. There the IDF scores for each language changed dynamically - for each new document we computed the IDF of all news articles within a 10 day window.
